### 2021-07-14 doc/thesis/chapters/implementation.tex
fixed point

\section{Digital Signal Processing Chain}

@@

\section{Matlab Model}

@@

\subsection{Transmitter}

@@

\subsection{Pre-Emphasis and De-Emphasis}
show effects in spectrum, show how to design the filters

\subsection{Fixed Point Arithmetic}
not developed in actual fixed point, but close to hardware level algorithm

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Digital Signal Processing Chain}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Matlab Model}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Transmitter}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pre-Emphasis and De-Emphasis}
TODO: show effects in spectrum, show how to design the filters

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fixed Point Arithmetic}

The target hardware platform of the FM radio receiver project is an FPGA.
These are devices that have a limited number of logic cells, which can be used to implement logical and computational functions.
Any kind of computation can be implemented within these logic cells.
However, depending on its detailed implementation, the amount of required logic cells may differ by a large factor.
In order to overcome this issue, mathematical operations are often implemented in fixed point arithmetic, rather than floating point.
This is due to the fact, that fixed point operations can be implemented with much less logic cells than floating point operations.
The reason therefor is, that the bitwidth of input and output numbers is previously known in fixed point.
Floating point calculations do not have this assumption and thus need to be more flexible, which requires a larger amount of logic cells.\\

The Matlab model is therefore developed by keeping in mind this limitation.
Matlab provides special datatypes for fixed point number representation.
However, these were not used in this implementation.
Instead, everything is calculated in floating point representation, using the standard \textit{double} datatype.
Then, in certain positions in the DSP chain, the numbers are rounded to the resolution of a fixed point datatype.
A fixed point format of \textit{2.14}, meaning 2 bit for integer- and 14 bit for fractional number representation is chosen.
This is sufficient for the FM receiver and could potentially be lowered to a lower bitwidth.
Again, an optimization of the DSP chain is not the focus of this thesis.\\

Summing up, the Matlab model is not implemented in fixed point arithmetic, but is implemented in a way, that is close enough to the hardware implementation.
Therefore, it can serve as a reference for the hardware implementation.
